---
title: "classnotes_b"
author: "Wesley Zuidema"
date: "March 31, 2016"
output: pdf_document
---

```{r}
library(dplyr)
library(uwpols501)
data(turnout)
data(iver)
```


Functions

```{r}
# Function example 1 - Use verbs for function names
verb <- function(argument1, argument2) {
  # do stuff
  print(argument1)
  print(argument2)
}
# Call verb
verb("ey", "oh")

# Functional example 2
say_hi <- function(name) {
  full_statement <- paste0("Hi! My name is ", name)
  return(full_statement)
}

say_hi("Wesley")

# Function example 3 - Get first digit
get_first_digit <- function(variable) {
  num_digits <- nchar(variable)
  var_first_num <- variable %/% (10 ^ (nchar(variable) - 1))
  return(var_first_num)
}

turnout$educate_new <- get_first_digit(turnout$educate)
turnout$age_new <- get_first_digit(turnout$age)

# Function example 4 - 
peripherial_countries <- c("Portugal", "Italy", "Ireland", 
                           "Cyprus", "Greece", "Spain")
create_per_dummy <- function(country_variable) {
  country_variable <- as.character(country_variable)
  dummy_boolean <- country_variable %in% peripherial_countries
  dummy_numeric <- as.numeric(dummy_boolean)
  return(dummy_numeric)
}

iver$per_dummy <- create_per_dummy(iver$cty)

# Function example 5 - own project
make_quantile_groups <- function(data, oldvar, newvar, prob) {
  data <- data %>%
    mutate(newvar = factor(as.numeric(
      oldvar > quantile(data$oldvar, probs = prob, 
                            na.rm = T))))
  data <- mutate(data, newvar = ifelse(newvar == 0, "low", "high"))
}
```

Loops

```{r}
for (each_element in here) { # Try to use meaningful names for the
  # do something             #   "each_element" object
  print(each_element)
}

list_numbers <- sample(x = 1:100, size = 10, replace = FALSE) # sample() function
for (number in list_numbers) {
  print(paste0("2 times ", number, " is ", number * 2))
}
```

Imagine that we want to calculate the log of the variables povred and enp of the iver dataset. We can create a list with these two variable names and then, for each of them, find them in iver, calculate the log, and add the result as a new variable to the dataset.

```{r}
numeric_vars <- c("povred", "enp")
for (variable in numeric_vars) {
  y <- iver[,variable]
  log_y <- log(y)
  name_new_var <- paste0(variable, "_log")
  iver[,name_new_var] <- log_y
}
```

Conditional loops

```{r}
if (this) {
  # do that
} else if (that) {
  # do something else
} else {
  #
}

some_numbers <- c(1, 4, 6, 10,12, 16, 45, 88, 102)
for (number in some_numbers) {
  if (number < 10) {
    print(paste0(number, " is smaller than 10"))
  } else if (number < 50) {
    print(paste0(number, " is smaller than 50 but greater or equal to 10"))
  } else {
    print(paste0(number, " is greater than 50"))
  }
}
```

We previously calculated the log and added it to the dataset for the numeric variables in iver. To do that we first created a list with the numeric variables. This time we won’t create that list. Instead, we’ll use conditional execution to indicate that we only want to take the log of numeric variables.

```{r}
data(iver) # load the data set again
iver <- as.data.frame(iver)
for (variable in names(iver)) {
  y <- iver[,variable]
  if (is.numeric(y)) {
    log_y <- log(y)
    name_new_var <- paste0(variable, "_log")
    iver[,name_new_var] <- log_y
  } else{
    print(paste0(variable, " is not a numeric variable."))
  }
}
```

